What We Eval in the Shadows

### 2021-04-16 paper/main.tex
Discussion Section 6

The examples are based on code from CRAN or Kaggle, simplified to fit the limited space of a scientific paper. We indicate the package and the function where the example is located.

@@

We group minimized expressions in larger categories.

@@

It is the predominant usage of \eval in Base, through function \c{match.args} (with \eval on minimized expression \c{V}). In CRAN, it correlates with more exotic environments passed to \eval. 
The dominant site in numbers of calls in CRAN, is located in \c{ggproto} in the custom object system of \emph{ggplot2}. It evaluates a variable \c{`\_inherit`}, which represents the parent of the class, in the definition environment of the class. \eval is used here to account for parents scoped in a namespace with \c{::}, which is actually a function call in R.

Package \emph{statnet.common} handles formulas, R object that associates an expression of the form \c{x~\textasciitilde~expr} and a carry-on environment where to evaluate the formula. It evaluates the left-hand side of the formula, which is always a variable, in function \c{eval\_lhs.formula}, as \c{eval(object[[2L]],envir=environment(object))} where \c{object} is the formula and \c{environment} returns the formula associated to the object.

@@

RunMcmc <- function... {
	 ...
	 eval(expr_q, envir=env)
\end{lstlisting}

@@

We present here high-level patterns that we identified after manual inspections of hundred of \eval sites. Some of the patterns were also described in the R documentation.

@@

Using \c{parse}, \ie converting strings to R expressions, opens to a more flexible code generation, as the building blocks which are going to be concatenated together, do not need to be valid R expressions. On the contrary, combining expression created with \c{substitute} or \c{quote} requires more discipline. In Kaggle, which features less experienced programmers,  code generation for plots is widespread, and uses \c{parse}.

@@

@@

\mypara{Discussion.} % more...

\eval is a powerful tool and can be sometimes replaced with functions with more specialized goals. For instance, variable lookup in any environment can be performed with \c{get}, and assignment in any environment, with \c{assign}. Building a call and executing can be done with \c{do.call}. Nevertheless, \eval shines when expressions passed to it can be of many different types. % And it is shorter to type, and no need to remember all the specialized functions.

>>>>

The examples are based on code from CRAN or Kaggle, simplified to fit the limited space of a scientific paper. For each example, we indicate the package and the function where it is located.

@@

We illustrate the minimized expressions of Section~\ref{sec:minimized} on representative examples.

% model.frame is in the next subsection, in the high-level patterns

@@

It is the predominant usage of \eval in Base, through function \c{match.args} (with evaluates on  minimized expression \c{V}). In CRAN, it correlates with more exotic environments passed to \eval. 
The dominant site in numbers of calls in CRAN, is located in \c{ggproto} in the custom object system of \emph{ggplot2}. It evaluates a variable \c{`\_inherit`}, which represents the parent of the class, in the definition environment of the class. \eval is used here to account for parents scoped in a namespace with \c{::}, which is actually a function call in R.

Package \emph{statnet.common} handles formulas, R object that associates an expression of the form \c{x~\textasciitilde~expr} and a carry-on environment where to evaluate the formula. It evaluates the left-hand side of the formula, which is always a variable, in function \c{eval\_lhs.formula}, where \c{object} is the formula and \c{environment} returns the formula associated to the object:

\begin{lstlisting}
eval(object[[2L]],envir=environment(object))
\end{lstlisting}

@@

RunMcmc <- function... {
	 eval(expr_q, envir=env)
\end{lstlisting}

@@

We present here high-level patterns that we identified after manual inspections of hundred of \eval sites. Some of the patterns are also described in the R documentation.

@@

Using \c{parse}, \ie converting strings to R expressions, opens to a more flexible code generation, as the building blocks which are going to be concatenated together, do not need to be valid R expressions. On the contrary, combining expression created with \c{substitute} or \c{quote} requires more discipline. In Kaggle, which features less experienced programmers,  code generation for plots is widespread, and uses \c{parse} on strings with \eval.

@@

\todo{Add a pattern about Library Loading? \c{library(packageName)} in R uses \eval This is one of the patterns in Javascript...}

@@

\mypara{Discussion.} \Eval in R is mostly used for meta-programming purposes and to access environments other than the local one. In Javascript, patterns are ad-hoc, such as json loading and parsing, and can often be rewritten without \eval, such as function or method call, or assignment to a local variable or an object property. Compared to Javascript, \eval patterns in R we observed are much more legitimate, although this is biased by the fact that our main corpus CRAN features more experienced developers than  websites found at large on the web. \todo{can we really claim that? SHould we talk about Kaggle there?}

\eval is a powerful tool and can be sometimes replaced with functions with more specialized goals. For instance, variable lookup in any environment can be performed with \c{get}, and assignment in any environment, with \c{assign}. Building a call and executing can be done with \c{do.call}. Nevertheless, \eval shines when expressions passed to it can be of many different types. % And it is shorter to type, and no need to remember all the specialized functions, so it is convenient.
